Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-05 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs:terse text on qeft

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/dcs_posteriors_combined.pdf}
\caption{(Color online).~Posterior distribution function for the coupling constant $\ell_{\rm CS}$ for
GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
%
In all panels, we use solid and dashed lines to correspond to the inclusion or not of the linear
in spin QNM correction.
%
We stress that posteriors obtained including $n=1$ corrections, violated conditions~\eqref{eq:parspec_bound}
and therefore should not be used to draw meaningful conclusions. We show them for illustrative purposes
and also to emphasize the importance of taking conditions~\eqref{eq:eft_bound} and~\eqref{eq:parspec_bound}
simultaneously into consideration when analyzing the results of the parameter estimation.
}
\label{fig:dCS_exec_sum}
\end{figure}

@@

We can draw qualitatively similar conclusions from the GW200129 event.
In particular, we find,
%
\begin{equation}
\ell_{\rm CS} \leqslant 36.6~\textrm{km}
% 43.9~\textrm{km}~( = \Lambda_{\rm EFT}(\tfrac{1}{2}, M_{f}) )
\quad \textrm{(90\% credibility)},
\end{equation}
%
and
%
\begin{equation}
\ell_{\rm CS} \leqslant 29.1~\textrm{km}
% ~( = \Lambda_{\rm EFT}(0.3, M_{f}) )
\quad \textrm{(68\% credibility)},
\end{equation}
%
when $N_{\rm max}=0$.
%
These stronger bound are a consequence of the larger support for $\ell_{\rm CS} \lessapprox 15$~km
for GW200129 (compare the top and middle panels in Fig.~\ref{fig:dCS_exec_sum}), and in part due to
the smaller median remnant ($M_{f} \approx 59.5\msun$ versus $M_{f} \approx 61.8\msun$ for GW150914).

@@

Finally, let us consider the quartic EFT of GR.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_posteriors_combined.pdf}
\caption{(Color online).~Posterior distribution function for the coupling constant $\ell_{\rm qEFT}$ for
GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
%
In all panels, we use solid and dashed lines to correspond to the inclusion or not of the linear
in spin QNM correction.
%
\hs{TODO}
}
\label{fig:qEFT_exec_sum}
\end{figure}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
              &       & bound? & bound?  &                            \\
\hline
  & GW150914 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 51.7$~km \\
0 & GW200129 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 54.8$~km \\
  & Combined &     &      & \cellcolor{black!10}$\ell_{\rm qEFT} \leqslant 51.3$~km \\
\hline
  & GW150914 &  &     &    \\
1 & GW200129 &  &     &    \\
  & Combined &  &     &    \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the quartic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
%
\hs{Need to debug $N_{\rm max} = 1$ runs.}
%
}
\label{tab:summary_qeft}
\end{table}

@@

We presented an unified framework to combine the framework introduced
to model deviations to the GR QNMs~\cite{Maselli:2019mjd} with the \pSEOB~waveform model~\cite{Brito:2018rfr,Ghosh:2021mrv}.
%
We showed with concrete examples how theory-specific calculations QNM of
slowly-rotating BHs in modified gravity theories can be mapped onto the
theory-agnostic beyond-GR parameters of the ParSpec formalism.
%
Put together this allowed us to use the \pSEOB~waveform model to test four
modified gravity theories (EdGB, dCS, cubic, and quartic EFTs of GR) using
observational data from the GW events GW150914 and GW200129.
%
We found, in particular that, within the assumptions of our model and by
stacking posteriors, that the coupling constant of dCS gravity is bound to
$\ell_{\rm CS} \leqslant 34.5$~km.
%
In contrast, we could place any constraint of the coupling constants in EdGB gravity $\ell_{\rm GB}$.
%
This dichotomy between these theories has an analogy with works that considered
the inspiral part of the GW signal alone.
%
In that case, it is dCS which is unconstrained, while EdGB, in principle can be
constrained, based on results for the closely related shift-symmetric version
of this theory.

>>>>

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/dcs_posteriors_combined.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:dCS_exec_sum}, but for dCS gravity.
%
% ~Posterior distribution function for the coupling constant $\ell_{\rm CS}$ for
% GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
% %
% In all panels, we use solid and dashed lines to correspond to the inclusion or not of the linear
% in spin QNM correction.
%
We stress that posteriors obtained including $n=1$ corrections, violated conditions~\eqref{eq:parspec_bound}
and therefore should not be used to draw meaningful conclusions. We show them for illustrative purposes
and also to emphasize the importance of taking conditions~\eqref{eq:eft_bound} and~\eqref{eq:parspec_bound}
simultaneously into consideration when analyzing the results of the parameter estimation.
}
\label{fig:dCS_exec_sum}
\end{figure}

@@

We can draw qualitatively similar conclusions from the GW200129 event.
In particular, we find,
%
\begin{equation}
\ell_{\rm CS} \leqslant 36.6~\textrm{km}
% 43.9~\textrm{km}~( = \Lambda_{\rm EFT}(\tfrac{1}{2}, M_{f}) )
\quad \textrm{(90\% credibility)},
\end{equation}
%
and
%
\begin{equation}
\ell_{\rm CS} \leqslant 29.1~\textrm{km}
% ~( = \Lambda_{\rm EFT}(0.3, M_{f}) )
\quad \textrm{(68\% credibility)},
\end{equation}
%
when $N_{\rm max}=0$.
%
These stronger bounds are a consequence of the larger support for $\ell_{\rm CS} \lessapprox 15$~km
for GW200129 (compare the top and middle panels in Fig.~\ref{fig:dCS_exec_sum}), and in part due to
the smaller median remnant ($M_{f} \approx 59.5\msun$ versus $M_{f} \approx 61.8\msun$ for GW150914).
%
\hs{Should we remove these and just quote the 90\% results?}

@@

At last, let us consider the quartic EFT of GR.
%
In Fig.~\ref{fig:qEFT_exec_sum} we show the posteriors on $\ell_{\rm qEFT}$ for GW150914 (top panel),
GW200129 (middle panel) for $N_{\rm max} = 0$, which are qualitatively similar to the cubic EFT of GR
%
We find that both the ``EFT''~\eqref{eq:eft_bound} and ``ParSpec''~\eqref{eq:parspec_bound} are satisfied,
hence allowing us to infer bounds on $\ell_{\rm qEFT}$.
%
As Fig.~\ref{fig:qEFT_cdf} shows the CDFs for both events reach 90\% for $\gm = M_f$ before $\varepsilon = 1/2$.
%
Consequently, from the joint posterior of $\ell_{\rm qEFT}$ we quote our final result,
%
\begin{equation}
    \ell_{\rm qEFT} \leqslant 51.3~\textrm{km}\,.
\end{equation}
%

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_posteriors_combined.pdf}
\caption{(Color online).~Posterior distribution function for the coupling constant $\ell_{\rm qEFT}$ for
GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
}
\label{fig:qEFT_exec_sum}
\end{figure}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
              &       & bound? & bound?  &                            \\
\hline
  & GW150914 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 51.7$~km \\
0 & GW200129 & Yes & Yes  & $\ell_{\rm qEFT} \leqslant 54.8$~km \\
  & Combined &     &      & \cellcolor{black!10}$\ell_{\rm qEFT} \leqslant 51.3$~km \\
\hline
  & GW150914 &  &     &    \\
1 & GW200129 &  &     &    \\
  & Combined &  &     &    \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the quartic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
}
\label{tab:summary_qeft}
\end{table}

@@

We presented an unified framework to combine the framework introduced
to model deviations to the GR QNMs~\cite{Maselli:2019mjd} with the \pSEOB~waveform model~\cite{Brito:2018rfr,Ghosh:2021mrv}.
%
We showed with concrete examples how theory-specific calculations QNM of
slowly-rotating BHs in modified gravity theories can be mapped onto the
theory-agnostic beyond-GR parameters of the ParSpec formalism.
%
Put together this allowed us to use the \pSEOB~waveform model to test four
modified gravity theories (EdGB, dCS, cubic, and quartic EFTs of GR) using
observational data from the GW events GW150914 and GW200129.
%
We found, in particular that, within the assumptions of our model and by
stacking posteriors, that the coupling constant of dCS gravity is bound to
$\ell_{\rm CS} \leqslant 34.5$~km.
%
In contrast, we could not place any constraint of the coupling constants in EdGB gravity $\ell_{\rm GB}$.
%
This dichotomy between these theories has an analogy with works that considered
the inspiral part of the GW signal alone.
%
In that case, it is dCS which is unconstrained, while EdGB, in principle can be
constrained, based on results for the closely related shift-symmetric version
of this theory.
